---
title: "LDL-C vs Death"
author: "Aljoscha Goetschi"
date: "`r date()`"
format: 
  html:
      toc: true
      code-fold: show
      code-tools: true
      df-print: paged
      embed-resources: true
---
```{r}
#| output: false
source("Packages.R")
```

## Introduction
In this short report, I am analysing whether the LDL-C concentration has an influence on whether patients die more, using the Framingham cohort study.

## Descriptive statistics
Three-thousand twenty-six LDL-C serum concentrations were available. The mean of LDL-C serum concentration amounted to 176.47 mg/dl (standard deviation = 46.86 mg/dl). The measurements ranged from 20 to 565 mg/dl. I calculated an exact 95% confidence interval according to Student of 174.80 - 178.14 mg/dl for the mean. 

```{r}
# Statistics

mean_ldl <- mean(framingham$LDLC, na.rm = T)
median_ldl <- median(framingham$LDLC, na.rm = T)
quartiles_ldl <- quantile(framingham$LDLC, na.rm = T, probs = c(0.25,0.5, 0.75))
var_ldl <- var(framingham$LDLC, na.rm = T)
sd_ldl <- sd(framingham$LDLC, na.rm = T)
range_ldl <- range(framingham$LDLC, na.rm = T)
iqr_ldl <- IQR(framingham$LDLC, na.rm = T)
# CIs
n <- length(which(!is.na(framingham$LDLC)))
tQuant <- qt(p = 0.975, df = n-1)
ldl_SE <- sd_ldl/sqrt(n)

CI_upper <- mean_ldl + tQuant*ldl_SE
CI_lower <- mean_ldl - tQuant*ldl_SE
```


```{r}
#| label: fig-dist
#| fig-cap: "Distribution of LDL-C serum concentrations. The mean is depcited in blue, and the median in red."
#| fig-width: 8
#| fig-height: 4
mean_ldl <- mean(framingham$LDLC, na.rm = T)
median_ldl <- median(framingham$LDLC, na.rm = T)

ggplot(framingham, aes(x = LDLC)) + geom_histogram(binwidth = 10) +
  geom_vline(xintercept = mean_ldl, colour = "blue") +
  geom_vline(xintercept = median_ldl, colour = "red") + 
  xlab(label = "LDL-C [mg/dl]") + ylab(label = "") + theme_bw()
```

